\Title

### 2020-11-09 manuscript/manuscript.tex

\subsection{Equivalent sources implementation}
\label{sec:eql-implementation}

@@

\subsection{Gradient boosted equivalent sources}

Reducing the number of sources helps, but it's not enough if we want to grid
a very large dataset.
Even if we can reduce the number of sources by one order of magnitude, having
too many data points would still increase the size of the Jacobian matrix until
the point of not being able to store it in memory.
This problem was already faced by other authors.

Here we propose an iterative algorithm for fitting the source coefficients
using a gradient boosting \citep{friedman2002} strategy and overlapping moving
windows.

Lets assume we have $N$ measurements of a harmonic field on observations points
$\{\mathbf{p}_i\}_{i=1}^{N}$ stored in the vector $\mathbf{d} = (d_1, d_2,
\ldots, d_N)^T$. We have defined a set of $M$ point sources located on
$\{\mathbf{q}_j\}_{j=1}^{M}$. Instead of trying to fit the entire set of
coefficients $\mathbf{c}$ at once, we can reduce the problem to fitting $K$
different sets of equivalent sources: we define $K$ overlapping windows of
equal size that cover the whole observation area, and then we create one set of
equivalent sources per window formed by the $\{\mathbf{q}_m\}_{m=1}^{M_k}$
sources that fall inside the $k$-th window.
The observed data should be equal to the superposition of the
effect of every set of equivalent sources:

\begin{equation}
    \mathbf{d} = \sum_{k=1}^K \mathbf{A}_k \mathbf{c}_k,
\end{equation}

\noindent where $\mathbf{c}_k$ is the vector of coefficients for the $k$-th
set of equivalent sources and $\mathbf{A}_k$ is the Jacobian matrix between
those sources and every data point.

By applying gradient boosting we can fit each $\mathbf{c}_k$ vector
iteratively: on each iteration a residue vector $\mathbf{r}_k$ (initialized
with the observed data) is used to fit the $\mathbf{c}_k$
coefficients and then updated by removing the effect of the same sources, as
follows:

\begin{equation}
    \mathbf{r}_0 = \mathbf{d},
\end{equation}

\begin{equation}
    \mathbf{r}_k = \mathbf{r}_{k-1} - \mathbf{A}_k\mathbf{c}_k.
    \label{eq:residue}
\end{equation}

\noindent On each iteration the $\mathbf{c}_k$ vectors can be fitted so they
minimize following misfit function:

\begin{equation}
    \phi(\mathbf{c}_k) =
        \left\lVert
            \mathbf{r}_{k-1} - \mathbf{A}_k\mathbf{c}_k
        \right\rVert ^ 2
        + \lambda_d \left\lVert \mathbf{c}_k \right\rVert ^2,
\end{equation}

\noindent following the same method described on Section
\ref{sec:eql-implementation}.
Because the $\mathbf{A}_k$ matrices have only $N \times M_k$ elements,
adjusting the size of the overlapping windows can help in reducing the computer
memory needed to fit the source coefficients and ultimately to interpolate the
observed data.

Nevertheless, in case our observed data is too large that even the
$\mathbf{A}_k$ matrices cannot be stored into memory, we can apply one more
simplification on the fitting process: ignore the effect of any observation
point outside the $k$-th window while fitting the sources that fall under the
same window.

Assuming that there are $N_k$ observation points that fall under the $k$-th
window, we can fit the $\mathbf{c}_k$ coefficients by minimizing the following
misfit function:

\begin{equation}
    \phi(\mathbf{c}_k) =
        \left\lVert
            \tilde{\mathbf{r}}_{k-1} - \tilde{\mathbf{A}}_k\mathbf{c}_k
        \right\rVert ^ 2
        + \lambda_d \left\lVert \mathbf{c}_k \right\rVert ^2,
\end{equation}

\noindent where $\tilde{\mathbf{r}}_{k-1}$ is the residue vector containing
only the elements that correspond to the $\{\mathbf{p}_n\}_{n=1}^{N_k}$
observation points that fall under the $k$-th window, and
$\tilde{\mathbf{A}}_k$ is the Jacobian matrix between the observation points
and sources that fall under the same window.
Because the $\tilde{\mathbf{A}}_k$ matrices will have $N_k \times M_k$
elements, the computer memory needed for fitting the $\mathbf{c}_k$
coefficients is greatly reduced.

Once the $\mathbf{c}_k$ coefficients are fitted, equation \ref{eq:residue} is
applied to update the residue, i.e. the effect of the sources belonging to the
$k$-th window is computed on every observation point. Worth noting that while
calculating the effect of the sources we don't need to pre-compute the entire
Jacobian matrix: the effect of the sources on each observation point can be
carried by a simple addition of terms, computing each Jacobian element at once
when needed.

Because the number of windows is fixed before starting the iteration process,
its number of steps are also fixed to $K$. Moreover, the windows are randomly
shuffled before starting the process to prevent that any particular
sets of equivalent sources are prioritized over the rest.

After all $\mathbf{c}_k$ coefficients vectors are fitted, we can predict the
effect of every equivalent source on any point as follows:

\begin{equation}
    V(\mathbf{p}) =
    \sum\limits_{k=1}^K \sum\limits_{m=1}^{M_k}
    \frac{{c_k}_m}{|\mathbf{p} - \mathbf{q}_m|}.
\end{equation}

>>>>

\subsection{Equivalent sources implementation}

@@
